---
title: "R graphics system series"
output: html_notebook
---

> Book **R Graphics**, Paul Murrell. The book code link is from [here](https://www.stat.auckland.ac.nz/~paul/RGraphics/rgraphics.html).

# Intro
simple example
```{r}
plot(pressure)+
text(150, 600, labels = "pressure (mm Hg) \nversus Temperature (Celsius)")
```
- In the R notebook, every plot will create newpage, so in base R graphics system, use `+` to run the code correctly
- The example is base R graphics in a nutshell, in order to produce graphical output, call a series of graphic function
- `R graphics` follow **painter models** that means the output occurs in steps with later output obscuring any previous output that it overlaps. R graphics fundamental is to add several graphical element to create final result.


```{r}

{
  # scatter plot
  par(mfrow=c(3, 2))
  x <- c(0.5, 2, 4, 8, 12, 16)
  y1 <- c(1, 1.3, 1.9, 3.4, 3.9, 4.8)
  y2 <- c(4, .8, .5, .45, .4, .3)
  par(las=1, mar=c(4, 4, 2, 4))
  plot.new()
  plot.window(range(x), c(0, 6))
  lines(x, y1)
  lines(x, y2)
  points(x, y1, pch=16, cex=2)
  points(x, y2, pch=21, bg="white", cex=2)
  par(col="grey50", fg="grey50", col.axis="grey50")
  axis(1, at=seq(0, 16, 4))
  axis(2, at=seq(0, 6, 2))
  axis(4, at=seq(0, 6, 2))
  box(bty="u")
  mtext("Travel Time (s)", side=1, line=2, cex=0.8)
  mtext("Responses per Travel", side=2, line=2, las=0, cex=0.8)
  mtext("Responses per Second", side=4, line=2, las=0, cex=0.8)
  text(4, 5, "Bird 131")
  par(mar=c(5.1, 4.1, 4.1, 2.1), col="black", fg="black", col.axis="black")
}
```












