---
title: "Foundations"
output: html_notebook
---

## Data structures

### Vectors

Basic data structure in R is the vector. Vectors come in two flavours: atomic & list.

1. Type, `typeof()`, what it is.
2. Length, `length()`, how many elements it contains.
3. Attributes, `attributes()`, additional arbitrary metadata.

Type of elements in atomic vector should be the same, whereas the elements of a list can have different types.

> *NB*: `is.vector()` does not test if an object is a vector. Instead it returns `TRUE` only if the object is a vector with no attributes apart from names. Use `is.atomic(x) || is.list(x)` to test if an object is actually a vector.

#### Atomic vectors

1. logical.
2. integer.
3. double.
4. charactor.

5. complex.
6. raw.
```{r}
fn_vector <- function(x){
  print(glue::glue("is.vector  {is.vector(x)}"))
  print(glue::glue("is.atomic {is.atomic(x)}"))
  print(glue::glue("is.list {is.list(x)}"))
  print(glue::glue("type is {typeof(x)}"))
  print(glue::glue("length is {length(x)}"))
  print(glue::glue("attributes are {attributes(x)}"))
}
dbl_var <- c(1, 2.5, 4.5)
fn_vector(dbl_var)
int_var <- c(1L, 2L, 3L)
fn_vector(int_var)
log_var <- c(T,F,T,F)
fn_vector(log_var)
chr_var <- c("these are", "some strings")
fn_vector(chr_var)
```

> `NA` is a logical vector of length 1. Creating `NA` of specific type with `NA_real_` (a double vector), `NA_integer` and `NA_character_`.

#### Types and tests

# Functional programming

## Functional programming

1. anonymous functions
2. closures
3. lists of functions
```{r}
set.seed(1014)
df <- data.frame(replicate(6, sample(c(1:10, -99), 6, rep = TRUE)))
names(df) <- letters[1:6]
df

fix_missing <- function(x) {
  x[x == -99] <- NA
  x
}
lapply(x = df, f = fix_missing)
out <- vector("list" ,length(df))
for ( i in seq_along(df)) {out[[1]] <- f(x[[i]], ...)}


fix_missing <- function(x) {
  x[x == -99] <- NA
  x
}

df[] <- lapply(df, fix_missing)

missing_fixer <- function(na_value) {
  function(x) {
    x[x == na_value] <- NA
    x
  }
}
fix_missing_99 <- missing_fixer(-99)
fix_missing_999 <- missing_fixer(-999)

fix_missing_99(c(-99, -999))
fix_missing_999(c(-99, -999))

# Storing function in a list
summa <- function(x) {
  funs <- c(mean, median, sd, mad, IQR)
  lapply(funs, function(f) f(x, na.rm = T))
}

summa(1:10)


```
### Anonymouse functions
```{r}
lapply(mtcars, function(x) length(unique(x)))
Filter(function(x) !is.numeric(x), mtcars)
integrate(function(x) sin(x) ^ 2, 0, pi)

formals(function(x = 4) g(x) + h(x))
body(function(x = 4) g(x) + h(x))
environment(function(x = 4) g(x) + h(x))


(function(x) x + 3)(10)
lapply(mtcars, function(x) mean(x) / sd(x))

integrate(function(x) x ^ 2 - x, 0, 10)
integrate(function(x) sin(x) + cos(x), -pi, pi)
```

### closures
```{r}
power <- function(e) function(x) x ^ e
power(2) %>% environment() %>% as.list()
power(3)(2)


new_counter <- function() {
  i <- 0
  function() {
    i <<- i + 1
    i
  }
}

counter_one <- new_counter()
counter_two <- new_counter()
counter_three <- new_counter()

counter_one()
counter_two()
counter_three()
```
### Lists of functions

```{r}
compute_mean <- list(
  base = function(x) mean(x),
  sum = function(x) sum(x) / length(x),
  manual = function(x) {
    total <- 0
    n <- length(x)
    for (i in seq_along(x)) {
      total <- total + x[i] / n
    }
    total
  }
)
compute_mean$base(runif(1e5)) 
x <- runif(1e5)
lapply(compute_mean, function(f) f(x))
call_fun <- function(f, ...) f(...)
lapply(compute_mean, call_fun, x)
lapply(compute_mean, function(f) system.time(f(x)))


simple_tag <- function(tag) {
  force(tag)
  function(...) {
    paste0("<", tag, ">", paste0(...), "</", tag, ">")
  }
}
tags <- c("p", "b", "i")
html <- lapply(setNames(tags, tags), simple_tag)

```




## Functionals
### lapply
```{r}
lapply2 <- function(x, f, ...){
  out <- vector("list", length(x))
  for (i in seq_along(x)) {
    out[[i]] <- f(x[[1]], ...)
  }
  out
}


l <- replicate(20, runif(sample(1:10, 1)), simplify = FALSE)


```

Looing patterns

1. loop over the elements: for (x in xs)
2. loop over the numeric indices: for (i in seq_along(xs))
3. loop over the names: for (name in names(xs))
```{r}
xs <- 1:10
res <- numeric(length(xs))
for (i in seq_along(xs)) {
  res[i] <- sqrt(xs[i])
}
str(res)

lapply(xs, function(x) {})
lapply(seq_along(xs), function(x) {})
lapply(names(xs), function(nm) {})

trims <- c(0, 0.1, 0.2, 0.5)
x <- rcauchy(100)

lapply(trims, function(trim) mean(x, trim = trim))

```


### Friends of lapply

1. `sapply()` and `vapply()`, variants of `lapply()` that produce vectors, matrices, and arrays as output, instead of list.
2. `Map()` and `mapply()` which iterate over multiple input data structure in parallel.
3. `mcapply()` and `mcMap()`, parallel versions of `lapply()` and `Map()`.

```{r}
sapply(mtcars, is.numeric)
vapply(mtcars, is.numeric, logical(1))
sapply(list(), is.numeric)
vapply(list(), is.numeric, logical(1))

df <- data.frame(x = 1:10, y = letters[1:10])
sapply(df, class)

vapply(df, class, FUN.VALUE = character(1))

df <- data.frame(x = 1:10, y = Sys.time() + 1:10)
sapply(df, class)

vapply(df, class, character(1))


sapply2 <- function(x, f, ...) {
  res <- lapply2(x, f, ...)
  simplify2array(res)
}

vapply2 <- function(x, f, f.value, ...) {
  out <- matrix(rep(f.value, length(x)), nrow = length(f.value))
  for (i in seq_along(x)) {
    res <- f(x[[i]], ...)
    stopifnot(
      length(res) == length(f.value),
      typeof(res) == typeof(f.value)
    )
    out[, i] <- res
  }
  out
}
vapply2(df[2], class, f.value = character(2))
```
```{r}
xs <- replicate(5, runif(10), simplify = F)
ws <- replicate(5, rpois(10, 5) + 1, simplify = F)

unlist(lapply(xs, mean))
unlist(lapply(seq_along(xs), function(i) {
  weighted.mean(xs[[i]], ws[[i]])
}))


unlist(Map(weighted.mean, xs, ws))
stopifnot(length(xs) == length(ws))
out <- vector("list", length(xs))
for (i in seq_along(xs)) {
  out[[i]] <- weighted.mean(xs[[i]], ws[[i]])
} 
out
```
```{r}
rollmean <- function(x, n) {
  out <- rep(NA, length(x))

  offset <- trunc(n / 2)
  for (i in (offset + 1):(length(x) - n + offset + 1)) {
    out[i] <- mean(x[(i - offset):(i + offset - 1)])
  }
  out
}
x <- seq(1, 3, length = 1e2) + runif(1e2)
plot(x)
lines(rollmean(x, 5), col = "blue", lwd = 2)
lines(rollmean(x, 10), col = "red", lwd = 2)

```
```{r}
boot_df <- function(x) x[sample(nrow(x), rep = T), ]
rsquared <- function(mod) summary(mod)$r.square
boot_lm <- function(i) {
  rsquared(lm(mpg ~ wt + disp, data = boot_df(mtcars)))
}

system.time(lapply(1:1000, boot_lm))
parallel::mclapply(1:1000, boot_lm, mc.cores = 2)
```
```{r}
integrate(sin, 0, pi)
uniroot(sin, pi * c(1 / 2, 3 / 2))
optimise(sin, c(0, 2 * pi), maximum = T)

```
```{r}
# create a function factory
# gives dataset, returns a function that computes the negative log likelihood for parameter lambda.
# closure
poisson_null <- function(x) {
  n <- length(x)
  sum_x <- sum(x)
  function(lambda) {
    n * lambda - sum_x * log(lambda)
  }
}

x1 <- c(41, 30, 31, 38, 29, 24, 30, 29, 31, 38)
x2 <- c(6, 4, 7, 3, 3, 7, 5, 2, 2, 7, 5, 4, 12, 6, 9)

nll1 <- poisson_null(x1)
nll2 <- poisson_null(x2)

optimise(nll1, c(0, 100))$minimum
optimise(nll2, c(0, 100))$minimum
```
```{r}
arg_max <- function(v, f, ...) {
  f(v) -> vf
  v[max(vf) == vf]
}
arg_max(-10:5, function(x) x ^ 2)
```

#### A family of functions
```{r}
add <- function(x, y) {
  stopifnot(
    length(x) == 1,
    length(y) == 1,
    is.numeric(x),
    is.numeric(y)
  )
  x + y
}

rm_na <- function(x, y, identity) {
  if (is.na(x) && is.na(y)) {
    identity
  } else if (is.na(x)){
    y
  } else {
    x
  }
}

add <- function(x, y, na.rm = F) {
  if (na.rm && (is.na(x) || is.na(y))) rm_na(x, y, 0) else x + y
}

r_add <- function(xs, na.rm = T) {
  Reduce(function(x, y) add(x, y, na.rm = na.rm), xs, init = 0)
}

r_add(c(1,4,10))
r_add(NA, na.rm = T)
r_add(numeric())


v_add1 <- function(x, y, na.rm = T) {
  stopifnot(
    length(x) == length(y),
    is.numeric(x), 
    is.numeric(y)
  )
  if (length(x) == 0) return(numeric())
  simplify2array(
    Map(function(x, y) add(x, y, na.rm = na.rm), x, y)
  )
}

v_add2 <- function(x, y, na.rm = T) {
  stopifnot(
    length(x) == length(y),
    is.numeric(x), 
    is.numeric(y)
  )
  if (length(x) == 0) return(numeric())
  
  vapply(
    X = seq_along(x),
    FUN = function(i) add(x[[i]], y[[i]], na.rm = na.rm),
    FUN.VALUE = numeric(1)
  )
}
v_add2(c(1,4,10), c(1,4,10))
```


## Function operators

```{r}
download_file <- function(url, ...) {
  download.file(url, basename(url), ...)
}
lapply(urls, download_file)

i <- 1
for (url in urls) {
  i <- i + 1
  if (i %% 10) cat(".")
  Sys.delay(1)
  download_file(url)
}

# use lapply like
lapply(urls, dot_every(10, delay_by(1, download_file)))

delay_by <- function(delay, f) {
  function(...) {
    Sys.sleep(delay)
    f(...)
  }
}

dot_every <- function(n, f) {
  i <- 1
  function(...) {
    if (i %% n == 0) cat(".")
    i <<- i + 1
    f(...)
  }
}
lapply(1:100, runif)
lapply(1:100, dot_every(10, runif))

download <- dot_every(10, delay_by(1, download.file))

```

```{r}
compact <- function(x) Filter(Negate(is.null), x)
```

```{r}
z <- quote(y <- x * 10)
str(z)
pryr::ast(y <- x * 10)
```

```{r}
subset2 <- function(x, condition) {
  contrib_call <- substitute(contition)
  r <- eval(contrib_call, x)
  x[r,]
}
```

# Expressions

- The structure of code is a tree.
- The branches of the tree are calls.
- The leaves of the tree are constants and names.
- Defin graphical conventions.


## AST (abstract syntax tree)
AST is usally constructed from a string, by parsing code. The opposite operation is deparsing. You can modify the tree by hand, and insert any R object. This can be useful when you need to override the usual lookup rules.

## Structure of expressions

- operation
- result

```{r}
z <- quote(y <- x * 10)
```










