---
title: "Kolmogorov-smirnov Test"
output: html_notebook
---
http://www.physics.csbsju.edu/stats/KS-test.html

> The Kolmogorov-Smirnov test (KS-test) tries to determine if two datasets differ significantly.

> The KS-test has the advantage of making no assumption about the distribution of data. (Technically speaking it is non-parametric and distribution free.)

> Note however, that this generality comes at some cost: other tests (for example Student's t-test) may be more sensitive if the data meet the requirements of the test.


```{r}
library(magrittr)
cA <- c(0.22, -0.87, -2.39, -1.79, 0.37, -1.54, 1.28, -0.31, -0.74, 1.72, 0.38, -0.17, -0.62, -1.10, 0.30, 0.15, 2.30, 0.19, -0.50, -0.09)
tA <- c(-5.13, -2.19, -2.43, -3.83, 0.50, -3.25, 4.32, 1.63, 5.18, -0.43, 7.11, 4.87, -3.10, -5.81, 3.76, 6.31, 2.58, 0.07, 5.76, 3.50)
t.test(tA, cA) %>% broom::tidy()
glue::glue("treatment mean: {mean(tA)} and sd: {sd(tA)}")
glue::glue("control mean: {mean(cA)} and sd: {sd(cA)}")


cB <- c(1.26, 0.34, 0.70, 1.75, 50.57, 1.55, 0.08, 0.42, 0.50, 3.20, 0.15, 0.49, 0.95, 0.24, 1.37, 0.17, 6.98, 0.10, 0.94, 0.38)
tB <- c(2.37, 2.16, 14.82, 1.73, 41.04, 0.23, 1.32, 2.91, 39.41, 0.11, 27.44, 4.51, 0.51, 4.50, 0.18, 14.68, 4.66, 1.30, 2.06, 1.19)
t.test(tB, cB)
glue::glue("treatment mean: {mean(tB)} and sd: {sd(tB)}")
glue::glue("control mean: {mean(cB)} and sd: {sd(cB)}")
ks.test(tB, cB) %>% broom::tidy()

```
```{r}
words1 = rnorm(100); words2 = rt(100, df=3)
plot(ecdf(words1))
## Have a look at the densities
plot(density(words1));plot(density(words2))

## Perform the test
shapiro.test(words1); shapiro.test(words2)

## Plot using a qqplot
qqnorm(words1);qqline(words1, col = 2)
qqnorm(words2);qqline(words2, col = 2)
```


```{r}
cBs <- cB %>% sort()
ecdf(cB) -> Fn
Fn(1)
sum(cB < 1) / length(cB)
plot(Fn, verticals = TRUE, do.points = FALSE)
plot(density(cB))
plot(density(tB))
shapiro.test(cB)
median(log(cB))
```

```{r}
plot(ecdf(log(cB)), verticals = TRUE, do.points = FALSE)
par(new = T)
plot(ecdf(log(tB)), verticals = TRUE, do.points = FALSE)

ks.test(cB, tB)
```

